---
title: "Low Level NHL API"
author: "Jozef Hajnala"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Low Level NHL API}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Low Level NHL API

If you would like to interact with the API directly without any helpers, `nhl_get_data()` function is made for this purpose. It takes as input one or more urls (a character vector) and returns a list of the same length, each element containing the raw results returned from the API for the corresponding URL.

For example:

```{r}
library(nhlapi)
urls <- c(
  "https://statsapi.web.nhl.com/api/v1/teams/1",
  "https://statsapi.web.nhl.com/api/v1/people/8477474"
)

nhl_get_data(urls)
```

Since it is helpful for work with R to get data in a flat data frame format, the function will by default flatten nested data frame columns and return a flat data frame. If you really want raw data returned, you can set the flatten argument to `FALSE`:

```{r}
nhl_get_data(urls, flatten = FALSE)
```

In case there are errors encountered when retrieving the data, the process does not fail, but returns a list with class `nhl_get_data_error`:

```{r}
nhl_get_data(c("https://statsapi.web.nhl.com/api/v1/wrongurl", urls))
```

The URLs themselves are preserved via the `url` attribute:

```{r}
lapply(nhl_get_data(urls), attr, which = "url")
```
