---
title: "Retrieving player data from the NHL API"
author: "Jozef Hajnala"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Retrieving player data from the NHL API}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# High-level players APIs

The high-level APIs intend to provide convenient functionality used by most users. Each function from the players API returns a data frame with the requested data processed to serve common purposes.

```{r}
library(nhlapi)
```

## Players information

The first endpoint retrieves information on players based on provided player names or ids. This information includes data such as names, nationalities, birth dates, etc.

One or more names of one or more ids can be provided to the function, data on each valid player name or id will be returned as one row of the resulting data frame:

With names

```{r}
# Get information on players by name
nhl_players(c("joe sakic", "PETER forsberg"))
```

With ids:

```{r}
playerIds <- c(8451101, 8458554)
nhl_players(playerIds = playerIds)
```

Invalid or erroring requests will not return any rows, the errors will be reported:

```{r}
nhl_players(c("made up player", "Joe Sakic"))

nhl_players(playerIds = c("made up player", 8451101))
```

The functions are designed is such a way that they allow working with the `%>%` operator if the user desires to do so:

```{r}
# Requires dplyr attached using library
nhl_players(playerIds) %>% 
  select(fullName, nationality, shootsCatches, primaryPosition.code)
```


## Players' season statistics

The second endpoint is used to retrieve players' statistics per seasons. The function `nhl_players_seasons` accepts multiple `playerIds` and multiple `seasons`. The proper API URLs will be constructed and a flat data frame returned:

```{r}
# Get a specific season statistics using player Ids
nhl_players_seasons(
  playerIds = c(8451101, 8458554),
  seasons = c("19961997", "19971998")
)
```

It also accepts a `playoff` argument in case the users wants playoff statistics returned and not regular season:

```{r}
# Get a specific season playoff statistics
nhl_players_seasons(
  playerIds = c(8451101, 8458554), 
  seasons = "19961997",
  playoffs = TRUE
)
```

A simple wrapper is also provided to get the statistics for all the seasons for the specified players:

```{r}
# Get all season statistics for a player
nhl_players_allseasons(playerIds = c(8451101, 8458554))
```

# Combining the endpoints

The package outputs are designed in such a way that they support easy data manipulation and exploration. For instance, using the popular `dplyr` package:

```{r}
# Requires dplyr attached using library
playerIds <- c(8451101, 8458554)

result <- nhl_players(playerIds) %>% 
  left_join(
    nhl_players_allseasons(playerIds),
    by = c("id" = "playerId")
  ) %>% 
  filter(league.name == "National Hockey League") %>% 
  select(fullName, seasonStart, stat.points)

head(result)
```

# Plotting the data

The nhlapi package itself does not contain plotting functionality, the output of user-facing functions are however designed to be data frames adhering to principles that make plotting with popular packages very natural.

For instance, using the ggplot2 package to plot the amount of points gathered by multiple players over the seasons:

```{r}
# Requires ggplot2 attached using library
ggplot(result) +
  geom_line(aes(seasonStart, stat.points, colour = fullName))
```

Or using the highcharter package to create an interactive version of the same chart:

```{r}
# Requires highcharter attached using library
result %>% 
  hchart("line", hcaes(seasonStart, stat.points, group = fullName)) %>%
  hc_add_theme(hc_theme_google())
```
