---
title: "Retrieving team data from the NHL API"
author: "Jozef Hajnala"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Retrieving team data from the NHL API}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# High-level team APIs

The high-level team APIs intend to provide convenient functionality used by most users. Each function from the players API returns a data frame with the requested data processed to serve common purposes.

```{r}
library(nhlapi)
```

## Team information

```{r}
# Get meta information on all teams
nhl_teams()
```

# Current rosters

```{r}
# Get current rosters for all teams
rosters <- nhl_teams_rosters()

# View the current roster for the Devils
roster_devils <- rosters %>%
  filter(name == "New Jersey Devils") %>%
  pull(roster.roster) %>%
  # This is still a list, extract the first element
  first()
roster_devils
```

# Rosters for past season

```{r}
# Get rosters for a specific season, all teams
rosters_1993 <- nhl_teams_rosters(seasons = 1993)

# Select the Quebec Nordiques
roster_nordiques_1993 <- rosters_1993 %>%
  filter(name == "Quebec Nordiques") %>%
  pull(roster.roster)
```

# Multiple seasons rosters

```{r}
# Works with multiple seasons
rosters_1993_1994 <- nhl_teams_rosters(
  seasons = c(1993:1995)
)

# It is a bit nicer though (for traffic) to get less data
# and make more specific requests - e.g. only get the Nordiques
rosters_1993_1994 <- nhl_teams_rosters(
  teamIds = 32,
  seasons = c(1993:1995)
)
```

# Upcoming and previous games

## Details on previous game for teams

```{r}
# Previous game for all teams
nhl_teams_shedule_previous()

# Previous game for selected teams
nhl_teams_shedule_previous(c(1,3,5))
```

## Exploring the full game details

There is a lot of information about games in the API, for instance, full details of the previous of the Devils:

```{r}
# First get the link
game_url <- nhl_teams_shedule_previous(1) %>%
  extract2("previousGameSchedule.dates") %>% first() %>%
  extract2("games") %>% first() %>%
  extract2("link")

# Now use `nhl_get_data()` to retrieve
game_data <- file.path("https://statsapi.web.nhl.com", game_url) %>%
  nhl_get_data()

# This is a very complex endpoint, for instance,
# look at the recorded events in the game
game_plays <- game_data %>% first() %>%
  extract2("liveData") %>%
  extract2("plays") %>%
  extract2("allPlays") %>%
  as_tibble()

game_plays
```


## Details on upcoming game for teams

```{r}
# Next game for all teams
nhl_teams_shedule_next()

# Next game for selected teams
nhl_teams_shedule_next(c(1,3,5))
```

# Team Stats per season

```{r}
# All teams, current seasons
nhl_teams_stats()
```

## Investigate the returned object

```{r}
# Get 2 seasons for the Devils:
devils_stats <- nhl_teams_stats(1, c("20052006", "20062007"))

# Look at the teamStats object for the first of them
devils_stats %>%
  extract2("teamStats") %>% first() %>%
  extract2("splits") %>% first() %>%
  slice(1) %>%
  select(stat.gamesPlayed, stat.wins, stat.losses)
```
